Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-07-12 doc/manuscript/manuscript.Rmd
fixed some issues; add Bruvo

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or white mold nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The white mold nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).

@@

Each isolate was regenerated by activating one or several sclerotia according to the method described previously [@matheron1989in-vitro].
Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
Since MCG are considered mutually exclusive, pairing was not required between all isolates in the study.

@@

Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center. 
Alleles were binned using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.

@@

Of the 16 microsatellite loci, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including samples that were potentially collected from the same plant (which increases the probability of collecting the same sample more than once), data were clone-corrected on the region/source/host/year hierarchy for subsequent analyses.

The clone-corrected data set included 318 isolates.
All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/

@@

![
Biplot showing 10 most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
Points represent the 318 clone-corrected samples in ordination space. 
](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=50%}

@@

Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
The three communities with the most members contained 8 of the ten most abundant MCGs. 
Overall, however, there was no clear pattern to the association between MLG and MCGs.

>>>>

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or white mold nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The white mold nurseries in different regions are indicated by wmn (white mold nursery) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).

@@

Each isolate was regenerated by activating one or several sclerotia according to the method described previously [@matheron1989in-vitro].
Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCGs resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
Since MCGs are considered mutually exclusive, pairing was not required between all isolates in the study.

@@

Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center. 
Alleles were scored using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.

@@

All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
<!-- add something about repeat length consistency  -->
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including samples that were potentially collected from the same plant (which increases the probability of collecting the same sample more than once), data were clone-corrected on the region/source/host/year hierarchy for subsequent analyses. The clone-corrected data set included 318 isolates.

@@

![
Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
Points represent the 318 clone-corrected samples in ordination space. 
](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=50%}

@@

Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
The three communities with the most members contained 8 of the ten most abundant MCGs. 
Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities. 
Overall, there was no clear pattern to the association between MLG and MCGs.
